### 2025-06-19 basic-statistics.qmd
20250619 - quantiles

- standard deviation ($s$ or $\sigma$)
- variance ($s^2$ or $\sigma^2$)
- range
- minimum and maximum (or some other quantiles)
- interquartile range (IQR)
- median absolute deviation
- coefficient of variation

@@

@@

```{r}
quantile(player_stats_seasonal$fantasyPoints, 0.10, na.rm = TRUE)
quantile(player_stats_seasonal$fantasyPoints, 0.90, na.rm = TRUE)
```

@@

player_stats_seasonal %>% 
  dplyr::select(age, years_of_experience, fantasyPoints) %>% 
  dplyr::summarise(across(
      everything(),
      .fns = list(
        n = ~ length(na.omit(.)),
        missingness = ~ mean(is.na(.)) * 100,
        M = ~ mean(., na.rm = TRUE),
        SD = ~ sd(., na.rm = TRUE),
        min = ~ min(., na.rm = TRUE),
        max = ~ max(., na.rm = TRUE),
        q10 = quantile(., .10, na.rm = TRUE), # 10th quantile
        q90 = quantile(., .90, na.rm = TRUE), # 90th quantile
        range = ~ max(., na.rm = TRUE) - min(., na.rm = TRUE),
        IQR = ~ IQR(., na.rm = TRUE),
        MAD = ~ mad(., na.rm = TRUE),
        CV = ~ sd(., na.rm = TRUE) / mean(., na.rm = TRUE),
        median = ~ median(., na.rm = TRUE),
        pseudomedian = ~ DescTools::HodgesLehmann(., na.rm = TRUE),
        mode = ~ petersenlab::Mode(., multipleModes = "mean"),
        skewness = ~ psych::skew(., na.rm = TRUE),
        kurtosis = ~ psych::kurtosi(., na.rm = TRUE)),
      .names = "{.col}.{.fn}")) %>%
    tidyr::pivot_longer(
      cols = everything(),
      names_to = c("variable","index"),
      names_sep = "\\.") %>% 
    tidyr::pivot_wider(
      names_from = index,
      values_from = value)
```

>>>>

- standard deviation ($s$ or $\sigma$)
- variance ($s^2$ or $\sigma^2$)
- range
- minimum and maximum (or some other quantiles)
- weighted quantiles
- interquartile range (IQR)
- median absolute deviation
- coefficient of variation

@@

If you want to give some values more weight to others, you can calculate weighted quantiles, while assigning a weight to each value.
The [`petersenlab`](https://github.com/DevPsyLab/petersenlab) package [@R-petersenlab] contains various functions for computing weighted quantiles based on @Akinshin2023.
Because some projections are outliers, we use a trimmed version of the weighted Harrell-Davis quantile estimator for greater robustness.

@@

```{r}
quantile(player_stats_seasonal$fantasyPoints, c(0.10, 0.90), na.rm = TRUE)
```

- weighted quantiles (in this case, the 10th and 90th quantiles):

```{r}
petersenlab::wthdquantile(
  player_stats_seasonal$fantasyPoints,
  w = sample( # randomly generate weights (could specify them manually)
    x = 1:3,
    size = length(player_stats_seasonal$fantasyPoints),
    replace = TRUE),
  probs = c(0.10, 0.90))
```

@@

player_stats_seasonal %>% 
  dplyr::select(age, years_of_experience, fantasyPoints) %>% 
  dplyr::summarise(across(
      everything(),
      .fns = list(
        n = ~ length(na.omit(.)),
        missingness = ~ mean(is.na(.)) * 100,
        M = ~ mean(., na.rm = TRUE),
        SD = ~ sd(., na.rm = TRUE),
        min = ~ min(., na.rm = TRUE),
        max = ~ max(., na.rm = TRUE),
        q10 = ~ quantile(., .10, na.rm = TRUE), # 10th quantile
        q90 = ~ quantile(., .90, na.rm = TRUE), # 90th quantile
        range = ~ max(., na.rm = TRUE) - min(., na.rm = TRUE),
        IQR = ~ IQR(., na.rm = TRUE),
        MAD = ~ mad(., na.rm = TRUE),
        CV = ~ sd(., na.rm = TRUE) / mean(., na.rm = TRUE),
        median = ~ median(., na.rm = TRUE),
        pseudomedian = ~ DescTools::HodgesLehmann(., na.rm = TRUE),
        mode = ~ petersenlab::Mode(., multipleModes = "mean"),
        skewness = ~ psych::skew(., na.rm = TRUE),
        kurtosis = ~ psych::kurtosi(., na.rm = TRUE)),
      .names = "{.col}.{.fn}")) %>%
    tidyr::pivot_longer(
      cols = everything(),
      names_to = c("variable","index"),
      names_sep = "\\.") %>% 
    tidyr::pivot_wider(
      names_from = index,
      values_from = value)
```

### 2025-06-19 correlation.qmd
20250619 - quantiles

```{r}
player_stats_seasonal %>% 
  dplyr::select(height, fantasyPoints) %>% 
  dplyr::summarise(across(
      everything(),
      .fns = list(
        n = ~ length(na.omit(.)),
        missingness = ~ mean(is.na(.)) * 100,
        M = ~ mean(., na.rm = TRUE),
        SD = ~ sd(., na.rm = TRUE),
        min = ~ min(., na.rm = TRUE),
        max = ~ max(., na.rm = TRUE),
        range = ~ max(., na.rm = TRUE) - min(., na.rm = TRUE),
        IQR = ~ IQR(., na.rm = TRUE),
        MAD = ~ mad(., na.rm = TRUE),
        CV = ~ sd(., na.rm = TRUE) / mean(., na.rm = TRUE),
        median = ~ median(., na.rm = TRUE),
        pseudomedian = ~ DescTools::HodgesLehmann(., na.rm = TRUE),
        mode = ~ petersenlab::Mode(., multipleModes = "mean"),
        skewness = ~ psych::skew(., na.rm = TRUE),
        kurtosis = ~ psych::kurtosi(., na.rm = TRUE)),
      .names = "{.col}.{.fn}")) %>%
    tidyr::pivot_longer(
      cols = everything(),
      names_to = c("variable","index"),
      names_sep = "\\.") %>% 
    tidyr::pivot_wider(
      names_from = index,
      values_from = value)
```

>>>>

```{r}
player_stats_seasonal %>% 
  dplyr::select(height, fantasyPoints) %>% 
  dplyr::summarise(across(
      everything(),
      .fns = list(
        n = ~ length(na.omit(.)),
        missingness = ~ mean(is.na(.)) * 100,
        M = ~ mean(., na.rm = TRUE),
        SD = ~ sd(., na.rm = TRUE),
        min = ~ min(., na.rm = TRUE),
        max = ~ max(., na.rm = TRUE),
        q10 = ~ quantile(., .10, na.rm = TRUE), # 10th quantile
        q90 = ~ quantile(., .90, na.rm = TRUE), # 90th quantile
        range = ~ max(., na.rm = TRUE) - min(., na.rm = TRUE),
        IQR = ~ IQR(., na.rm = TRUE),
        MAD = ~ mad(., na.rm = TRUE),
        CV = ~ sd(., na.rm = TRUE) / mean(., na.rm = TRUE),
        median = ~ median(., na.rm = TRUE),
        pseudomedian = ~ DescTools::HodgesLehmann(., na.rm = TRUE),
        mode = ~ petersenlab::Mode(., multipleModes = "mean"),
        skewness = ~ psych::skew(., na.rm = TRUE),
        kurtosis = ~ psych::kurtosi(., na.rm = TRUE)),
      .names = "{.col}.{.fn}")) %>%
    tidyr::pivot_longer(
      cols = everything(),
      names_to = c("variable","index"),
      names_sep = "\\.") %>% 
    tidyr::pivot_wider(
      names_from = index,
      values_from = value)
```

### 2025-06-19 multiple-regression.qmd
20250619 - quantiles

```{r}
player_stats_seasonal %>% 
  dplyr::select(fantasyPoints, age, height, weight, target_share) %>% 
  dplyr::summarise(across(
      everything(),
      .fns = list(
        n = ~ length(na.omit(.)),
        missingness = ~ mean(is.na(.)) * 100,
        M = ~ mean(., na.rm = TRUE),
        SD = ~ sd(., na.rm = TRUE),
        min = ~ min(., na.rm = TRUE),
        max = ~ max(., na.rm = TRUE),
        range = ~ max(., na.rm = TRUE) - min(., na.rm = TRUE),
        IQR = ~ IQR(., na.rm = TRUE),
        MAD = ~ mad(., na.rm = TRUE),
        median = ~ median(., na.rm = TRUE),
        pseudomedian = ~ DescTools::HodgesLehmann(., na.rm = TRUE),
        mode = ~ petersenlab::Mode(., multipleModes = "mean"),
        skewness = ~ psych::skew(., na.rm = TRUE),
        kurtosis = ~ psych::kurtosi(., na.rm = TRUE)),
      .names = "{.col}.{.fn}")) %>%
    tidyr::pivot_longer(
      cols = everything(),
      names_to = c("variable","index"),
      names_sep = "\\.") %>% 
    tidyr::pivot_wider(
      names_from = index,
      values_from = value)
```

>>>>

```{r}
player_stats_seasonal %>% 
  dplyr::select(fantasyPoints, age, height, weight, target_share) %>% 
  dplyr::summarise(across(
      everything(),
      .fns = list(
        n = ~ length(na.omit(.)),
        missingness = ~ mean(is.na(.)) * 100,
        M = ~ mean(., na.rm = TRUE),
        SD = ~ sd(., na.rm = TRUE),
        min = ~ min(., na.rm = TRUE),
        max = ~ max(., na.rm = TRUE),
        q10 = ~ quantile(., .10, na.rm = TRUE), # 10th quantile
        q90 = ~ quantile(., .90, na.rm = TRUE), # 90th quantile
        range = ~ max(., na.rm = TRUE) - min(., na.rm = TRUE),
        IQR = ~ IQR(., na.rm = TRUE),
        MAD = ~ mad(., na.rm = TRUE),
        median = ~ median(., na.rm = TRUE),
        pseudomedian = ~ DescTools::HodgesLehmann(., na.rm = TRUE),
        mode = ~ petersenlab::Mode(., multipleModes = "mean"),
        skewness = ~ psych::skew(., na.rm = TRUE),
        kurtosis = ~ psych::kurtosi(., na.rm = TRUE)),
      .names = "{.col}.{.fn}")) %>%
    tidyr::pivot_longer(
      cols = everything(),
      names_to = c("variable","index"),
      names_sep = "\\.") %>% 
    tidyr::pivot_wider(
      names_from = index,
      values_from = value)
```
